Modal type theory based on the intuitionistic epistemic logic

### 2018-01-14 article.tex
added rewrite

  \begin{defin} Intuitionistic normal modal logic {\bf IK}

    2) $\Box (A \to B) \to (\Box A \to \Box B)$;

    3) Rules: MP and necessiation.

  \begin{defin} Translation from {\bf IK} into {\bf IEL}$^{-}$

    1) $\bot^{\circ} = \bot$;

    2) $p^{\circ} = p$;

    3) $(A \alpha B)^{\circ} = A^{\circ} \alpha B^{\circ}$, $\alpha \in \{ \to, \wedge, \vee \}$.

    4) $(\Box A)^{\circ} = {\bf K} A^{\circ}$
  \end{defin}

  \begin{lemma}
    $ $
    ${\bf IK} \vdash A \Rightarrow {\bf IEL}^{-} \vdash A^{\circ}$
  \end{lemma}

  \begin{proof}
    Straightforward induction on the structure of $M$.
  \end{proof}

  It is clearly that we may prove similar fact for ${\bf IEL}$ by the same way.

@@

  	\begin{prooftree}
    	\AxiomC{$\Gamma \vdash A$}
    	\RightLabel{$\textbf{K}_I$}
    	\UnaryInfC{$\Gamma \vdash \textbf{K}A$}
  \end{prooftree}
    \end{flushleft}
  \end{minipage}
  \begin{minipage}{0.5\textwidth}
    \begin{flushright}
    	\begin{prooftree}
  	\AxiomC{$\Gamma \vdash A \to B$}
  	\AxiomC{$\Gamma \vdash A$}
  	\RightLabel{$\rightarrow_i$}
  	\BinaryInfC{$\Gamma \vdash B$}
    	\end{prooftree}

@@

  	\begin{prooftree}
    \AxiomC{$\Gamma \vdash {\bf K} \vec{A}$}
    \AxiomC{$\vec{A} \vdash B$}
    \BinaryInfC{$\Gamma \vdash {\bf K} B$}
  	\end{prooftree}
    \end{flushright}
  \end{minipage}
  \end{defin}

Where $\Gamma \vdash {\bf K} \vec{A}$ is a syntax sugar for $\Gamma \vdash {\bf K} A_1 ,\dots, \Gamma \vdash {\bf K} A_n$.

@@

1) If $\Gamma \vdash_{NIEL^{-}_{\land, \to}} A$, then $IEL^{-} \vdash \bigwedge \Gamma \rightarrow \textbf{K}A$.

$\begin{array}{lll}
(1) & \bigwedge \Gamma \rightarrow A & \text{assumption}\\
(2) & A \rightarrow \textbf{K}A &\text{co-reflection}\\
(3) & (\bigwedge \Gamma \rightarrow A) \rightarrow ((A \rightarrow \textbf{K}A) \rightarrow (\bigwedge \Gamma \rightarrow \textbf{K}A))&\text{IPC theorem}\\
(4) & (A \rightarrow \textbf{K}A) \rightarrow (\bigwedge \Gamma \rightarrow \textbf{K}A) &\text{from (1), (3) and MP}\\
(5) & \bigwedge \Gamma \rightarrow \textbf{K}A &\text{from (2), (4) and MP}\\
\end{array}$

@@

2) If $\Gamma \vdash_{NIEL^{-}_{\land, \to}} {\bf K} \vec{A}$ and $\vec{A} \vdash B$, then $IEL^{-} \vdash \bigwedge \Gamma \rightarrow \textbf{K}B$.

$\begin{array}{lll}
(1) &\bigwedge \Gamma \rightarrow \bigwedge \limits_{i = 1}^{n} {\bf K} A_i & \text{assumption} \\
(2) &\bigwedge \limits_{i = 1}^{n} {\bf K} A_i \rightarrow {\bf K} \bigwedge \limits_{i = 1}^{n} A_i& \text{IEL theorem} \\
(3) &\bigwedge \Gamma \rightarrow {\bf K} \bigwedge \limits_{i = 1}^{n} A_i & \text{from (1), (2) and transitivity} \\
(4) &\bigwedge \limits_{i = 1}^{n} A_i \rightarrow B& \text{assumption} \\
(5) &(\bigwedge \limits_{i = 1}^{n} A_i \rightarrow B) \rightarrow {\bf K} (\bigwedge \limits_{i = 1}^{n} A_i \rightarrow B)& \text{co-reflection}\\
(6) &{\bf K} (\bigwedge \limits_{i = 1}^{n} A_i \rightarrow B)& \text{from (2), (3) and MP} \\
(7) &{\bf K} \bigwedge \limits_{i = 1}^{n} A_i \rightarrow {\bf K} B & \text{from (6) and normality} \\
(8) &\bigwedge \Gamma \rightarrow {\bf K} B & \text{from (3), (7) and transitivity}
\end{array}$

@@

@@

@@

@@

$\begin{array}{lll}
&\text{1) } (\lambda x. M) N \rightarrow_{\beta} M [x := N]& \\
&\text{2) } \pi_1 \langle M, N \rangle \rightarrow_{\beta} M& \\
&\text{3) } \pi_2 \langle M, N \rangle \rightarrow_{\beta} N& \\
&\text{4) } {\bf let \: pure \: } \vec{y} = \vec{M_2} {\bf \: in \: } ({\bf let \: pure \: } \vec{x} = \vec{M_1} {\bf \: in \: }  N) \rightarrow_{\beta}& \\
&\quad {\bf let \: pure \: } \vec{x} = ({\bf let \: pure \: } \vec{y} = \vec{M_2} \: {\bf in} \: \vec{M_1}) \: {\bf in} \: {\bf let \: pure \: } \vec{x} = \vec{x} {\bf \: in \:} N&
\end{array}$

@@

  So $\Gamma \uparrow FV(\vec{M}) \vdash {\bf let \: pure \:} \vec{x} = \vec{M} {\bf \: in \:} N : \textbf{K}B$.

@@

  2) .

@@

@@

  We consider cases with reduction rules which are applicative laws. The general statement for $\twoheadrightarrow_{\beta}$ follows from transitivity of multi-step $\beta$-reduction.

@@

  1) Let $\Gamma \vdash {\bf pure \: } (\lambda x. x) \star M : \textbf{K}\alpha$. Then $\Gamma \vdash
  {\bf pure \: } (\lambda x.x) : \textbf{K}(\alpha \to \alpha)$ and $\Gamma \vdash M : \textbf{K}\alpha$ by
  generation. Then $\Gamma \vdash M : \textbf{K}\alpha$ trivially.

  2) Let $\Gamma \vdash {\bf pure \: }(\lambda f g x. f (g x)) \star M \star N \star P : \textbf{K}\gamma$.

  Then $\Gamma \vdash {\bf pure \: } (\lambda f g x. f (g x)) : \textbf{K}((\beta \to \gamma) \to (\alpha \to \beta)
  \to \alpha \to \gamma)$, $\Gamma \vdash M : \textbf{K}(\beta \to \gamma)$, $\Gamma \vdash N :
  \textbf{K}(\alpha \to \beta)$ and $\Gamma \vdash P : \textbf{K}\alpha$ by generation.

  If $\Gamma \vdash N : \textbf{K}(\alpha \to \beta)$ and $\Gamma \vdash P : \textbf{K}\alpha$, then $\Gamma
  \vdash N \star P : \textbf{K}\beta$ by $\textbf{K}_{app}$.

  Hence, if $\Gamma \vdash M : \textbf{K}(\beta \to \gamma)$, then $\Gamma \vdash M \star (N \star P) :
  \textbf{K}\gamma$ by $\textbf{K}_{app}$.

  3) Let $\Gamma \vdash ({\bf pure \: } M) \star ({\bf pure \: } N) : \textbf{K}\beta$. Then $\Gamma \vdash
  {\bf pure \: } M : \textbf{K}(\alpha \to \beta)$ and $\Gamma \vdash {\bf pure \: } N : \textbf{K}\alpha$ by
  generation. Moreover, $\Gamma \vdash M : \alpha \to \beta$ and $\Gamma \vdash N : \alpha$.

  Then $\Gamma \vdash M N : \beta$ by application.

  Hence, $\Gamma \vdash {\bf pure \: } (M N) : \textbf{K}\beta$ by $K_I$.

  4) Let $\Gamma \vdash M \star ({\bf pure \: } N) : \textbf{K}\beta$.

  Then $\Gamma \vdash M : \textbf{K}(\alpha \to \beta)$ and $\Gamma \vdash {\bf pure \: } N : \textbf{K}\alpha$.

  Moreover, $\Gamma \vdash N : \alpha$ by generation.

  Let $\Gamma, f : \alpha \to \beta \vdash f : \alpha \to \beta$ and $\Gamma, f : \alpha \to \beta \vdash N :
  \alpha$ by weakening.

  So $\Gamma, f : \alpha \to \beta \vdash f N : \beta$ by application, so $\Gamma \vdash \lambda f. f N :
  (\alpha \to \beta) \to \beta$ by abstraction.

  Then $\Gamma \vdash {\bf pure \: } (\lambda f. f N) : \textbf{K}((\alpha \to \beta) \to \beta)$ by $\textbf{K}_I$.

  Hence, $\Gamma \vdash {\bf pure \: } (\lambda f. f N) \star M : \textbf{K}\beta$.

@@

  \bibitem{Baren2} H.P. Barendregt. The Lambda Calculus --- Its Syntax and Semantics. Studies in Logic and
  the Foundations of Mathematics, vol. 103. Amsterdam: North-Holland, 1985.
  \end{thebibliography}

>>>>

@@

  	\begin{prooftree}
    	\AxiomC{$\Gamma, \Delta \vdash A$}
    	\RightLabel{$\vdash {\bf K}_I$}
    	\UnaryInfC{${\bf K}(\Gamma), \Delta \vdash {\bf K}A$}
  \end{prooftree}
    \end{flushleft}
  \end{minipage}
  \begin{minipage}{0.5\textwidth}
    \begin{flushright}
    	\begin{prooftree}
  	\AxiomC{$\Gamma \vdash A \to B$}
  	\AxiomC{$\Gamma \vdash A$}
  	\RightLabel{$\rightarrow_i$}
  	\BinaryInfC{$\Gamma \vdash B$}
    	\end{prooftree}

@@

  	\begin{prooftree}
    \AxiomC{${\bf K}(\Gamma), \Delta \vdash {\bf K} A$}
    \AxiomC{${\bf K}(\Gamma), {\bf K} A, \Delta \vdash B$}
    \BinaryInfC{${\bf K}(\Gamma), \Delta \vdash B$}
  	\end{prooftree}
    \end{flushright}
  \end{minipage}
  \end{defin}

@@

1) If $\Gamma, \Delta \vdash_{NIEL^{-}_{\land, \to}} A$, then $IEL^{-} \vdash \bigwedge {\bf K} \Gamma \wedge \bigwedge \Delta \rightarrow {\bf K}A$.

See \cite{Krupski}.

@@

2) If ${\bf K} \Gamma, \Delta \vdash_{NIEL^{-}_{\land, \to}} {\bf K} A$ and ${\bf K}\Gamma, {\bf K} A, \Delta \vdash B$, then $IEL^{-} \vdash \bigwedge {\bf K} \Gamma \wedge \bigwedge \Delta \rightarrow B$.

$\begin{array}{lll}
&\text{(1) } \bigwedge {\bf K} \Gamma \wedge \bigwedge \Delta \to {\bf K} A& \text{assumption} \\
&\text{(2) } \bigwedge {\bf K} \Gamma \wedge {\bf K}A \wedge \bigwedge \Delta \rightarrow B& \text{assumption} \\
&\text{(3) } {\bf K} A \wedge (\bigwedge {\bf K} \Gamma \wedge \bigwedge \Delta) \rightarrow B& \text{conjuction properties} \\
&\text{(4) } ({\bf K} A \wedge (\bigwedge {\bf K} \Gamma \wedge \bigwedge \Delta) \rightarrow B) \rightarrow ({\bf K } A \rightarrow ((\bigwedge {\bf K} \Gamma \wedge \bigwedge \Delta) \rightarrow B)) & \text{IPC theorem} \\
&\text{(5) } {\bf K} A \rightarrow ((\bigwedge {\bf K} \Gamma \wedge \bigwedge \Delta) \rightarrow B) & \text{MP } \\
&\text{(6) } \bigwedge {\bf K} \Gamma \wedge \bigwedge \Delta \to ((\bigwedge {\bf K} \Gamma \wedge \bigwedge \Delta) \rightarrow B) & \text{(1), (5), transitivity} \\
&\text{(7) } (\bigwedge {\bf K} \Gamma \wedge \bigwedge \Delta) \rightarrow B & \text{(6), contraction}
\end{array}$

@@

  \begin{lemma}
    $ $

    If $\text{IEL}^{-} \vdash A$, then $\text{NIEL}^{-} \vdash A$.
  \end{lemma}

  \begin{proof}
    $ $
    Straightforward derivation of modal axioms in NIEL.
  \end{proof}

@@

  We divide into two parts, the first one zone is a modal part (enviroment with applicaitve types), the second one
  is a intuitionistic part (enviroment with pure types). From a logical point of view, $\Gamma | \Delta \vdash M : A$ denotes that
  $M$ is a proof of ${\bf K} \Gamma, \Delta \vdash A$. This idea similar to \cite{DePaiva}.

@@

  2) $(M N) [x := P] = M[x := P] N [x := P]$;

@@

$\begin{array}{lll}
&\text{1) } (\lambda x. M) N \rightarrow_{\beta} M [x := N]& \\
&\text{2) } \pi_1 \langle M, N \rangle \rightarrow_{\beta} M& \\
&\text{3) } \pi_2 \langle M, N \rangle \rightarrow_{\beta} N& \\
\end{array}$

@@

  So $\Gamma \uparrow FV(\vec{M}) \vdash {\bf let \: pure \:} \vec{x} = \vec{M} {\bf \: in \:} N : \textbf{K}B$ by ${\bf let}_{\textbf{K}}$.

@@

  2) Let $\Gamma, y : C \vdash {\bf let \: pure \:} \vec{x} = \vec{M} {\bf \: in \:} N : \textbf{K}B$ and $\Gamma \vdash P : C$.

  By generation $\Gamma, y : C \vdash \vec{M} : \textbf{K}\vec{A}$ and $\vec{x} : \vec{A} \vdash N : \textbf{K}B$.

  By assumption $\Gamma \vdash \overrightarrow{M [y := P]} : \textbf{K}A$.

  So $\Gamma \vdash {\bf let \: pure \:} \vec{x} = \overrightarrow{M [y := P]} {\bf \: in \:} N : \textbf{K}B$ by ${\bf let}_{\textbf{K}}$.

@@

  \begin{defin} Type substitution

    The substitution of type $C$ for type variable $B$ in some type $A$ is defined by:

    1) $A [ A := B ] = B$;

    2) $A^{'} [A := C] = A^{'}$;

    3) $(A_1 \alpha A_2) [B := C] = A_1 [B := C] \alpha A_2 [A := B]$, where $\alpha \{ \times, \to \}$;

    4) $(\textbf{K}A) [B := C] = \textbf{K}(A [B := C])$.

  \begin{lemma} Substitution for types.

    Let $\Gamma \vdash M : A$, then $\Gamma [B := C] \vdash M : A [B := C]$, where $\Gamma [B := C]$ denotes $\{ x_1 : A_1 [B := C],\dots, x_n : A_n [B := C] \}$.

  1) Let $\Gamma, x : A \vdash x : A$, so $\Gamma [B := C], x : A [B := C] \vdash x : A [B := C]$.

  2) Let $\Gamma \vdash {\bf pure \: } M : \textbf{K}A$. By generation $\Gamma \vdash M : A$.

  By assumption $\Gamma [B := C] \vdash M : A [B := C]$.

  By $\textbf{K}_I$, $\Gamma [B := C] \vdash {\bf pure \:} M : \textbf{K} (A [B := C])$.

  But, by substitution definition, $\textbf{K} (A [B := C]) = (\textbf{K}A) [B := C]$.

  Then $\Gamma [B := C] \vdash {\bf pure \:} M : (\textbf{K}A) [B := C]$.

  3) $\Gamma \vdash {\bf let \: pure} \: \vec{x} = \vec{M} \: {\bf in} \: M : {\bf K} B$.

  $\Gamma [D := C] \vdash \vec{M} : \textbf{K}\overrightarrow{A [D := C]}$ and $\vec{x} : \overrightarrow{A [D := C]} \vdash M : B [D := C]$ by assumption.

  Then $\Gamma [D := C] \vdash {\bf let \: pure } \: \vec{x} = \vec{M} \: {\bf in} \: M : {\bf } (B [D := C])$.

@@

@@

@@

  \bibitem{Baren2} H.P. Barendregt. The Lambda Calculus --- Its Syntax and Semantics. Studies in Logic and
  the Foundations of Mathematics, vol. 103. Amsterdam: North-Holland, 1985.

  \bibitem{DePaiva} de Paiva V., Ritter E., Bellin, G. Extended Curry-Howard correspondence for a basic constructive modal logic. Methods for Modalities Conference (M4M-2); 2001 November 29-30; Amsterdam; the Netherlands.
\end{thebibliography}
